---
title: "622 GROUP PROJECT 3"
author: "Jessica Valencia"
date: "2023-03-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=FALSE}
# Tidyverse and text analysis tools
library(tidyverse)
library(tidytext)

# For decision trees
library(rpart)
library(rpart.plot)
library(rattle)

# For nearest neighbors
library(class)

# For ML evaluation
library(caret)
```

First, we brought in the data file after handcoded sentiment has been conducted.
```{r}
library(readr)
datasubset <- read_csv("hw3datasubset.csv")
#View(hw3datasubset)
```
#REMOVING NEUTRAL SENTIMENT FROM HANDCODED SENTIMENT##############


Then, we removed the observations with a "neutral" hand-coded sentiment from the data set considering that this information is not useful for the purpose of this report. 
```{r}
data<-datasubset[datasubset$Handcoded_Sentiment!="Neutral",]

View(data)
```

#CLEANING HANDCODED SENTIMENT##############


We decided to clean the hand-coded sentiment column in order to make sure our results are as optimal and accurate as possible by removing human error. 
```{r}
data$Handcoded_Sentiment <- replace(data$Handcoded_Sentiment, data$Handcoded_Sentiment == "Posotive", "Positive")

data$Handcoded_Sentiment <- replace(data$Handcoded_Sentiment, data$Handcoded_Sentiment == "Postive", "Positive")

data$Handcoded_Sentiment <- replace(data$Handcoded_Sentiment, data$Handcoded_Sentiment == "postive", "Positive")
```



